Reliable error estimation for Sobol' indices\thanks{This work is supported by the CITiES project funded by the Agence Nationale de la Recherche (grant ANR-12-MONU-0020) and by the United States National Science Foundation (grant DMS-1522687).}

### 2017-05-21 article_review2.0/Manuscript.tex
added two changes

\subsection{Estimation of Sobol' indices}
\label{sec:2.2}
In this section we review two Monte Carlo procedures for the estimation of Sobol' indices. A design is a point set $\mathcal{P}=\{\vx_i\}_{i=0}^{n-1}$ where each point is obtained by sampling each variable $x_j$ $n$ times. Each row of the design is a point $\vx_i$ in $[0,1]^d$ and each column of the design refers to samples of a variable $x_j$. Consider $\mathcal{P}=\{\vx_i\}_{i=0}^{n-1}$ and $\mathcal{P'}=\{{\vx'}_i\}_{i=0}^{n-1}$ two designs where $(\vx_i,{\vx'}_i) \iid \mathcal{U}([0,1]^{2d})$. One way to estimate the two quantities (\ref{first.order}) and (\ref{total.effect}) is via:
\begin{align}
\label{first.order.est}
\widehat{\underline{\tau}_u^2} &= \frac{1}{n} \sum \limits_{i=0}^{n-1} \left(f(\vx_{i,u}:{\vx'}_{i,-u})-f(\vx'_i)\right)f(\vx_i),\\
\label{total.effect.est}
\widehat{\ov{$\tau$}_u^2} &= \frac{1}{2n} \sum \limits_{i=0}^{n-1} (f({\vx'}_i) - f(\vx_{i,u}:{\vx'}_{i,-u}))^2, \qquad u \in \mathcal{D},
\end{align}
using for $\sigma^2$:
\begin{equation}\label{mu.est}
\widehat{\sigma}^2 = \frac{1}{n-1} \sum \limits_{i=0}^{n-1} \left(f(\vx_i) - \widehat{\mu}\right)^2, \text{ with } \ \widehat{\mu} =  \frac{1}{n} \sum \limits_{i=0}^{n-1}f(\vx_i).
\end{equation}
\comment{Equation \eqref{mu.est} can be improved if we use the $2n$ evaluations $f(\vx_i)$ and $f(\vx'_i)$. While this would be an important improvement to estimate the indices' denominator, the error construction described in Section \ref{sec:3} would require additional computations. Because of the new computations in the algorithm, the additional benefit of using $2n$ evaluations could become more costly than beneficial.}

Then, the Sobol' indices estimators are:
\begin{equation}
\widehat{\underline{S}}_u = \widehat{\underline{\tau}_u^2} / \widehat{\sigma}^2, \qquad
\widehat{\overline{S}}_u = \widehat{\overline{\tau}_u^2} / \widehat{\sigma}^2.
\label{common.sobol.est}
\end{equation}
Using this method, the estimation of a single pair ($\widehat{\underline{S}}_u$, $\widehat{\overline{S}}_u$) requires $3n$ evaluations of the model $f$. Thus, for all first order and total effect indices, one would need $3nd$ model evaluations. Using a combinatorial formalism, in \cite{Saltelli} Saltelli proposes the following estimation strategy:
\begin{theorem}
\label{saltelli.theorem}
The $d+2$ designs $\{\vx_{i,u}:{\vx'}_{i,-u}\}_{i=0}^{n-1}$ constructed for $u \in \{\varnothing,\{1\},\dots,$ $\{d\},\mathcal{D}\}$ allow to estimate all first-order and total effect Sobol' indices at a cost of $n(d+2)$ evaluations of the model.
\end{theorem}
The main idea of the theorem is that there is no need to reevaluate $f({\vx}_i)$ and $f({\vx'}_i)$ for each $u$. Hence, we can simply evaluate $f({\vx}_i)$ $n$ times, $f({\vx'}_i)$ $n$ times, and $f(\vx_{i,u}:{\vx'}_{i,-u})$ $nd$ times, which accounts for the $n(d+2)$ evaluations.

@@

\begin{figure}[t]
\caption{Boxplots of estimation errors $\delta_{\underline{S}_u}$ obtained with both Variant $A.b$ and the iid sampling method. The dashed horizontal line marks the tolerance $\varepsilon=5\cdot 10^{-3}$.}
\centering
\vspace*{0.2cm}
\includegraphics[width=0.4\textwidth]{Fig5.eps}
\label{compa_MC_first}
\end{figure}

\begin{figure}[t]
\caption{Boxplots of estimation errors $\delta_{\overline{S}_u}$ obtained with both Variant $A.b$ and the iid sampling method. The dashed horizontal line marks the tolerance $\varepsilon=5\cdot 10^{-3}$.}
\centering
\vspace*{0.2cm}
\includegraphics[width=0.4\textwidth]{Fig6.eps}
\label{compa_MC_tot}
\end{figure}

>>>>

\subsection{Estimation of Sobol' indices}
\label{sec:2.2}
In this section we review two Monte Carlo procedures for the estimation of Sobol' indices. A design is a point set $\mathcal{P}=\{\vx_i\}_{i=0}^{n-1}$ where each point is obtained by sampling each variable $x_j$ $n$ times. Each row of the design is a point $\vx_i$ in $[0,1]^d$ and each column of the design refers to samples of a variable $x_j$. Consider $\mathcal{P}=\{\vx_i\}_{i=0}^{n-1}$ and $\mathcal{P'}=\{{\vx'}_i\}_{i=0}^{n-1}$ two designs where $(\vx_i,{\vx'}_i) \iid \mathcal{U}([0,1]^{2d})$. One way to estimate the two quantities (\ref{first.order}) and (\ref{total.effect}) is via:
\begin{align}
\label{first.order.est}
\widehat{\underline{\tau}_u^2} &= \frac{1}{n} \sum \limits_{i=0}^{n-1} \left(f(\vx_{i,u}:{\vx'}_{i,-u})-f(\vx'_i)\right)f(\vx_i),\\
\label{total.effect.est}
\widehat{\ov{$\tau$}_u^2} &= \frac{1}{2n} \sum \limits_{i=0}^{n-1} (f({\vx'}_i) - f(\vx_{i,u}:{\vx'}_{i,-u}))^2, \qquad u \in \mathcal{D},
\end{align}
using for $\sigma^2$:
\begin{equation}\label{mu.est}
\widehat{\sigma}^2 = \frac{1}{n-1} \sum \limits_{i=0}^{n-1} \left(f(\vx_i) - \widehat{\mu}\right)^2, \text{ with } \ \widehat{\mu} =  \frac{1}{n} \sum \limits_{i=0}^{n-1}f(\vx_i).
\end{equation}
\comment{Equation \eqref{mu.est} can be improved if we use the $2n$ evaluations $f(\vx_i)$ and $f(\vx'_i)$. While this would be an important improvement to estimate the indices' denominator, the error construction described in Section \ref{sec:3} would require additional computations. Because of these new computations in the algorithm, the additional benefit of using $2n$ evaluations could become more costly than beneficial.}

The Sobol' indices estimators are:
\begin{equation}
\widehat{\underline{S}}_u = \widehat{\underline{\tau}_u^2} / \widehat{\sigma}^2, \qquad
\widehat{\overline{S}}_u = \widehat{\overline{\tau}_u^2} / \widehat{\sigma}^2.
\label{common.sobol.est}
\end{equation}
Based on equation \eqref{common.sobol.est}, the estimation of a single pair ($\widehat{\underline{S}}_u$, $\widehat{\overline{S}}_u$) requires $3n$ evaluations of the model $f$. Thus, for all first order and total effect indices, one would need $3nd$ model evaluations. Using a combinatorial formalism, in \cite{Saltelli} Saltelli proposes the following estimation strategy:
\begin{theorem}
\label{saltelli.theorem}
The $d+2$ designs $\{\vx_{i,u}:{\vx'}_{i,-u}\}_{i=0}^{n-1}$ constructed for $u \in \{\varnothing,\{1\},\dots,$ $\{d\},\mathcal{D}\}$ allow to estimate all first-order and total effect Sobol' indices at a cost of $n(d+2)$ evaluations of the model.
\end{theorem}
The main idea of the theorem is that there is no need to reevaluate $f({\vx}_i)$ and $f({\vx'}_i)$ for each $u$. Hence, we can simply evaluate $f({\vx}_i)$ $n$ times, $f({\vx'}_i)$ $n$ times, and $f(\vx_{i,u}:{\vx'}_{i,-u})$ $nd$ times, which accounts for the $n(d+2)$ evaluations.

@@

\begin{figure}[t]
\caption{Boxplots of estimation errors $\delta_{\underline{S}_u}$ obtained with both Variant $A.b$ and the iid sampling method. The dashed horizontal line marks the tolerance $\varepsilon=5\cdot 10^{-3}$.}
\centering
\vspace*{0.2cm}
\includegraphics[width=0.45\textwidth]{Fig5.eps}
\label{compa_MC_first}
\end{figure}

\begin{figure}[t]
\caption{Boxplots of estimation errors $\delta_{\overline{S}_u}$ obtained with both Variant $A.b$ and the iid sampling method. The dashed horizontal line marks the tolerance $\varepsilon=5\cdot 10^{-3}$.}
\centering
\vspace*{0.2cm}
\includegraphics[width=0.45\textwidth]{Fig6.eps}
\label{compa_MC_tot}
\end{figure}
